---
output: github_document
title: "NGCHM R package"
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# NGCHM R package

<!-- badges: start -->
<!-- badges: end -->

NGCHM is an R package for creating [Next-Generation Clustered Heat Maps](https://bioinformatics.mdanderson.org/public-software/ngchm/).
Next-Generation Clustered Heat Maps (NG-CHMs) allow for a dynamic exploration of heat map data in a web browser. 

Our [YouTube Video](https://www.youtube.com/watch?v=DuObpGNpDhw) presents a brief overview of NGCHM usage and functionality.

Click [Start Here!](articles/create-a-ng-chm.html) in the main menu to learn how to create an NG-CHM like the one below.


### NG-CHM of Breast Cancer (BRCA) Data from The Cancer Genome Atlas (TCGA)

```{r, message=FALSE, warning=FALSE, echo=FALSE }
library(NGCHMDemoData)
library(NGCHMSupportFiles)
library(NGCHM)
unadjustedColorMap <- chmNewColorMap(c(6.4,10,14), c('mediumblue','snow','firebrick'))
rowCenteredColorMap <- chmNewColorMap(c(-2,0,2), c('#9933ff','#f0f0f0','#228B22'))
unadjustedLayer <- chmNewDataLayer('Unadjusted', TCGA.BRCA.ExpressionData, unadjustedColorMap)
rowCenteredData <- t(scale(t(TCGA.BRCA.ExpressionData)))
rowCenteredLayer <- chmNewDataLayer('Row-Centered', rowCenteredData, rowCenteredColorMap)
hm <- chmNew('TCGA BRCA Expression', unadjustedLayer, rowCenteredLayer)
covariateBar <- chmNewCovariate('TP53 Mutation',TCGA.BRCA.TP53MutationData)
hm <- chmAddCovariateBar(hm, 'column', covariateBar)
hm <- chmAddAxisType(hm, 'row', 'bio.gene.hugo')
hm <- chmAddAxisType(hm, 'column', 'bio.tcga.barcode.sample.vial.portion.analyte.aliquot')
```

```{r, message=FALSE, warning=FALSE, echo=FALSE, results='hide' }
chmExportToHTML(hm,'tcga-brca-intro.html',overwrite=TRUE)
```

```{r, message=FALSE, warning=FALSE, echo=FALSE }
library('htmltools')
filePath = paste(getwd(),'/tcga-brca-intro.html',sep='')
htmltools::includeHTML(filePath)
```


