---
title: "Clustering and Column/Row Ordering"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Clustering and Column/Row Ordering}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette describes options for ordering rows and columns of a 
NG-CHM, and the additional options available when the ordering is hierarchical clustering.

## Specifying Row / Column Order

The `chmNew()` function allows for arguments "rowOrder" and "colOrder", each of which can be a vector, dendrogram, or function
specifying the order of the rows/columns.

The following command produces a NG-CHM with rows sorted alphabetically and columns ordered randomly:

```{r, results='hide', message=FALSE, warning=FALSE, eval=FALSE}
library(NGCHM)
library(NGCHMDemoData)
hm <- chmNew('tcga-gbm', TCGA.GBM.ExpressionData, rowOrder=sort(rownames(TCGA.GBM.ExpressionData)), colOrder=colnames(TCGA.GBM.ExpressionData)[ sample.int(length(colnames(TCGA.GBM.ExpressionData))) ])
```

## Hierarchical Clustering Options

For hierarchical clustering, additional options are available to specify the distance measure and clustering method. 

The distance measure is specified with the 'rowDist' and 'colDist' arguments to `chmNew()`, and 
available choices include those of the `stats::dist()` function. 
There are two additional distance metric options: 'cosine' and 'correlation'. The default is 'correlation', 
which computes the distance measure as 1 minus the Pearson correlation among the rows/columns.

The clustering method is specified with the 'rowAgglom' and 'colAgglom' arguments, and
available options are those of the `stats::hclust` function. The default is 'ward.D2'.

This example specifies a NG-CHM with hierarchical clustering using the Euclidean distance 
metric and complete linkage clustering algorithm for both rows and columns:

```{r, results='hide', message=FALSE, warning=FALSE, eval=FALSE}
hm <- chmNew('tcga-gbm', TCGA.GBM.ExpressionData, rowDist='euclidean', colDist='euclidean', rowAgglom='complete', colAgglom='complete')
```

## Explicit Clustering

Users may perform the clustering explicitly, and use the results in constructing a NG-CHM.
The following creates an object of class hclust from the distance matrix of the demo 
expression data, for both the rows and columns,
and uses those clustering results to construct an NG-CHM.

```{r, results='hide', message=FALSE, warning=FALSE, eval=FALSE}
rowClust <- stats::hclust(dist(TCGA.GBM.ExpressionData))
colClust <- stats::hclust(dist(t(TCGA.GBM.ExpressionData)))
hm <- chmNew('tcga-gbm', TCGA.GBM.ExpressionData, rowOrder=rowClust, colOrder=colClust)
```

Similarly, an hclust object can be transformed into a dendrogram and used for the row and/or column
ordering. The example below uses the clutering results from above as dedrograms:

```{r, results='hide', message=FALSE, warning=FALSE, eval=FALSE}
rowDendrogram <- as.dendrogram(rowClust)
colDendrogram <- as.dendrogram(colClust)
hm <- chmNew('tcga-gbm', TCGA.GBM.ExpressionData, rowOrder=rowDendrogram, colOrder=colDendrogram)
```

