---
title: "Continuous Color Maps and Data Layers"
output: rmarkdown::html_vignette
resource_files:
  - images/layersOff.png
vignette: >
  %\VignetteIndexEntry{Continuous Color Maps and Data Layers}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette demonstrates how to specify the color maps and use data layers by creating a 
NG-CHM from 2 data layers: a layer from the unadjusted data, and another from row-centering that data.
Each layer will have its own custom color map.

## Color Map (Continuous)

Color maps are created with `chmNewColorMap()`. The first argument to this function
is a list of breakpoints. For the TCGA.BRCA.ExpressionData demo data, reasonable breakpoints are
6.4, 10, and 14. The second argument is a corresponding list of
colors for those breakpoints. The colors can be specified by name for standard colors, or by hexadecimal code.

For the unadjusted data color map:

```{r, results='hide', message=FALSE, warning=FALSE}
library(NGCHM)
unadjustedColorMap <- chmNewColorMap(c(6.4,10,14), c('mediumblue','snow','firebrick'))
```

For the second color map to be used with row-centered data, breakpoints are set at -2, 0, and 2. For 
the purpose of illustration, the hexidecimal code is used for specifying the color.

```{r, results='hide', message=FALSE, warning=FALSE}
rowCenteredColorMap <- chmNewColorMap(c(-2,0,2), c('#9933ff','#f0f0f0','#228B22'))
```


## Data Layers 

Data layers can be created with the `chmNewDataLayer()` function. The first argument is the desired name 
of the data layer. The second argument is the matrix of data, and the third argument is the color map.

```{r, results='hide', message=FALSE, warning=FALSE}
library(NGCHMDemoData)
unadjustedLayer <- chmNewDataLayer('Unadjusted', TCGA.BRCA.ExpressionData, unadjustedColorMap)
rowCenteredData <- t(scale(t(TCGA.BRCA.ExpressionData)))
rowCenteredLayer <- chmNewDataLayer('Row-Centered', rowCenteredData, rowCenteredColorMap)
```

## Creating the NG-CHM

We explicitly provide the two data layers created above to the `chmNew()` function:

```{r, results='hide', message=FALSE, warning=FALSE}
hm <- chmNew('TCGA BRCA Expression', unadjustedLayer, rowCenteredLayer)
```

and export to a .ngchm or .html file:

```{r, results='hide', message=FALSE, warning=FALSE}
library(NGCHMSupportFiles)
chmExportToFile(hm,'tcga-brca-two-layers.ngchm', overwrite=TRUE)
chmExportToHTML(hm,'tcga-brca-two-layers.html', overwrite=TRUE)
```

## Resulting NG-CHM

<!-- svg symbol symbol taken from NG-CHM viewer:
     https://github.com/MD-Anderson-Bioinformatics/NG-CHM/blob/main/NGCHM/WebContent/icons.svg?short_path=fffa0d9
-->

<svg style="display:none;">
  <symbol id="icon-layers" xmlns="http://www.w3.org/2000/svg" fill="currentColor" class="bi bi-layers" viewBox="0 0 16 16">
    <path d="M8.235 1.559a.5.5 0 0 0-.47 0l-7.5 4a.5.5 0 0 0 0 .882L3.188 8 .264 9.559a.5.5 0 0 0 0 .882l7.5 4a.5.5 0 0 0 .47 0l7.5-4a.5.5 0 0 0 0-.882L12.813 8l2.922-1.559a.5.5 0 0 0 0-.882l-7.5-4zm3.515 7.008L14.438 10 8 13.433 1.562 10 4.25 8.567l3.515 1.874a.5.5 0 0 0 .47 0l3.515-1.874zM8 9.433 1.562 6 8 2.567 14.438 6 8 9.433z"/>
  </symbol>
</svg>

The resulting NG-CHM has the two data layers available. Click on the
<svg style="width:20px;height:20px;" viewBox="0 0 20 20">
    <use xlink:href="#icon-layers"></use>
</svg> icon (the first of the 4 buttons on the upper right side) to toggle between the data layers.

```{r, message=FALSE}
library('htmltools')
filePath = paste(getwd(),'/tcga-brca-two-layers.html',sep='')
htmltools::tags$iframe(src=filePath, width='100%', height=700)
```

[Back to top](#)

