---
title: "Covariate Bars and Discrete Color Maps"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Covariate Bars and Discrete Color Maps}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette demonstrates how to add covariate bars and use discrete color maps in a NG-CHM, and uses the 
sample data described in [Create a NG-CHM](create-a-ng-chm.html).

## Create a Covariate Bar 

To start, the demo data, support files, and NGCHM-R package are loaded into the current R environment, and 
an initial NGCHM of the [demo expression data](create-a-ng-chm.html#sample-ng-chm-data-1) is created with the `chmNew()` function.

```{r, results='hide', message=FALSE, warning=FALSE}
library(NGCHMDemoData)
library(NGCHM)
hm <- chmNew('tcga-brca', TCGA.BRCA.ExpressionData)
```

A covariate bar can be created from the TP53 mutation vector using the `chmNewCovariate()` function.
The first argument is the desired name of the covariate bar, and the second is the vector
of TP53 mutation demo data.

```{r, results='hide', message=FALSE, warning=FALSE}
covariateBar <- chmNewCovariate('TP53 Mutation',TCGA.BRCA.TP53MutationData)
```

This covariate bar is then added to the NG-CHM created above with `chmAddCovariateBar()`:

```{r, results='hide', message=FALSE, warning=FALSE}
hm <- chmAddCovariateBar(hm, 'column', covariateBar)
```

This NG-CHM can be exported to either a .ngchm or .html file.

```{r, results='hide', message=FALSE, warning=FALSE, eval=FALSE}
library(NGCHMSupportFiles)
chmExportToFile(hm,'tcga-brca-covariates.ngchm')
chmExportToHTML(hm,'tcga-brca-covariates.html')
```

## Specify Covariate Bar Color Map

In the section above, the color map for the 'TP53 Mutation' covariate bar was not specified, and the 
NGCHM-R package generated one automatically. However a color map can be explicitly constructed, as shown 
below.

The TP53 mutation data values consist of two discrete strings: 'WT' denoting wild type, and 'MUT' denoting
mutation. 

The `chmNewColorMap()` function is used for creating a color map, and a color map for the TP53 mutation data,
which consists of two strings: 'WT' and 'MUT' can be created with explicit color names or hexidecimal codes:


```{r, results='hide', message=FALSE, warning=FALSE}
mutationColorMap <- chmNewColorMap(c("WT","MUT"),c("cornflowerblue","lightcoral"))
## Or equivalently, because #6495ED and #f08080 are the hexidecimal color codes 
## for cornflowerblue and lightcoral, respectively:
mutationColorMap <- chmNewColorMap(c("WT","MUT"),c("#6495ED","#f08080"))
```

In the `chmNewCovariate()` function (also used above), an optional third argument is a color map:

```{r, results='hide', message=FALSE, warning=FALSE}
covariateBar <- chmNewCovariate('TP53 Mutation (other colors)',TCGA.BRCA.TP53MutationData,mutationColorMap)
```

Finally, this covariate bar can be added to the NG-CHM created above,

```{r, results='hide', message=FALSE, warning=FALSE}
hm <- chmAddCovariateBar(hm,'column',covariateBar)
```

As described in the [Create a NG-CHM](create-a-ng-chm.html) page, this NG-CHM
can be exported to either a .ngchm or .html file.

```{r, results='hide', message=FALSE, warning=FALSE}
library(NGCHMSupportFiles)
chmExportToFile(hm,'tcga-brca-covariates.ngchm')
chmExportToHTML(hm,'tcga-brca-covariates.html')
```


## Resulting NG-CHM

The resulting NG-CHM has both the original covariate bar with the default color map and the covariate bar (of the same data)
with the cornflowerblue and lightcoral color map:

```{r, echo = FALSE}
library('htmltools')
filePath = paste(getwd(),'/tcga-brca-covariates.html',sep='')
htmltools::tags$iframe(src=filePath, width='100%', height=700)
```

