---
title: "getting-started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{getting-started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Next Generation Clustered Heat Maps

For a complete overview of Next-Generation Clustered Heat Maps,
please see [our web page](https://bioinformatics.mdanderson.org/public-software/ngchm/).

This vignette demonstrates how to construct a simple NG-CHM and 
save it as a file that can be opened in the [NGCHM Viewer](http://www.ngchm.net/Downloads/ngChmApp.html).

# Sample NG-CHM Data and .jar file

This vignette requires two additional packages:

1. [NGCHMDemoData](https://github.com/MD-Anderson-Bioinformatics/NGCHMDemoData), a package of demo data, and 
2. [ShaidyMapGenJar](https://github.com/MD-Anderson-Bioinformatics/ShaidyMapGenJar), a package containing a Java 
   .jar file needed for constructing the final NGCHM.

These can be installed from GitHub with [devtools](https://cran.r-project.org/web/packages/devtools/index.html):

```{r message=FALSE, results='hide'}
library(devtools)
devtools::install_github('MD-Anderson-Bioinformatics/NGCHMDemoData')
devtools::install_github('MD-Anderson-Bioinformatics/ShaidyMapGenJar')
```

and loaded into the current environment:

```{r message=FALSE, results='hide'}
library(NGCHMDemoData)
library(ShaidyMapGenJar)
```

# Creating an NG-CHM

The sample data loaded from [NGCHMDemoData](https://github.com/MD-Anderson-Bioinformatics/NGCHMDemoData) 
includes a matrix of mRNA expression data, TCGA.GBM.Expression, for 3540 genes (rows) and 169 samples 
(columns) from the Glioblastoma Multiforme TCGA data.

An NGCHM, hm, of this data can be constructed with the chmNew() function:

```{r}
library(NGCHM)
hm <- chmNew('tcga-gbm', TCGA.GBM.ExpressionData)
```

This NGCHM can be exported to a file with the chmExportToFile() function.
**NOTE: The filename must end with '.ngchm'**

```{r message=FALSE, results='hide'}
chmExportToFile(hm,'tcga-gbm.ngchm')
```

This file can be opened in the [NGCHM Viewer](http://www.ngchm.net/Downloads/ngChmApp.html).

# Adding Covariate Bars

The [NGCHMDemoData](https://github.com/MD-Anderson-Bioinformatics/NGCHMDemoData) package includes
a vector of TP53 mutation data, TCGA.GBM.TP53MutationData, for the samples (columns) of the NGCHM constructed above.

For a vector to be used as an NGCHM covariate bar, the names of the vector should be 
identical to the colnames of the matrix data. In other words, for this example the following should 
evaluate as TRUE for each entry:

```{r message=FALSE, results='hide'}
names(TCGA.GBM.TP53MutationData) == colnames(TCGA.GBM.ExpressionData) 
```

The covariate bar can be created and added to the NGCHM with:

```{r}
covariateBar <- chmNewCovariate('TP53 Mutation',TCGA.GBM.TP53MutationData)
hm <- chmAddCovariateBar(hm, 'column', covariateBar)
```

The NGCHM can be exported to a file, overwriting the file we created above with,

```{r message=FALSE, results='hide'}
chmExportToFile(hm,'tcga-gbm.ngchm',overwrite=TRUE)
```

This file can be opened in the [NGCHM Viewer](http://www.ngchm.net/Downloads/ngChmApp.html).

