---
title: "Installation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Installation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## R Packages Installation

The NG-CHM R package can be installed from CRAN:

```{r, results='hide', message=FALSE, warning=FALSE, eval=FALSE}
install.packages("NGCHM")
```

The required NGCHMSupportFiles can be installed from
[our R-Universe repository](https://md-anderson-bioinformatics.r-universe.dev/packages):


```{r, results='hide', message=FALSE, warning=FALSE, eval=FALSE}
install.packages("NGCHMSupportFiles", repos = c("https://md-anderson-bioinformatics.r-universe.dev"))
```

The optional NGCHMDemoData (used in the vignette examples) can similarly be installed:

```{r, results='hide', message=FALSE, warning=FALSE, eval=FALSE}
install.packages("NGCHMDemoData", repos = c("https://md-anderson-bioinformatics.r-universe.dev"))
```

## System Requirements

- R >= 3.5
- Java >= 11
- git
  - Mac/Linux: typically installed by default
  - Windows: Cygwin or Windows Subsystem for Linux (Windows 10)
- C compiler
  - Mac: Xcode
  - Linux: gcc
  - Windows: Rtools or Rtools40 (Windows 10)
- [NGCHMSupportFiles](https://md-anderson-bioinformatics.r-universe.dev/NGCHMSupportFiles) R package
  - This package requires Java >= 11
  - This package contains a compiled .jar file and minified JavaScript file
- *Optional:* The following are only required if using a remote NG-CHM server:
  - tar, ssh, scp *(most users will not need these)*


## Docker Image

A Docker image containing RStudio, the NGCHM-R package, and all needed dependencies is available on Docker Hub:
[ngchm/rstudio-ngchm](https://hub.docker.com/r/ngchm/rstudio-ngchm).

Detailed usage information for the docker image is available
in this the [RStudio NGCHM GitHub](https://github.com/MD-Anderson-Bioinformatics/rstudio-ngchm) project,
and the YouTube Video,
[How to Create Next-Generation Clustered Heat Maps in R Studio](https://www.youtube.com/watch?v=O42w5P3A1_8).


