---
title: "Installation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Installation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## System Requirements

- R >= 3.5
- Java >= 11
- git, tar, scp, ssh
  - Mac/Linux: typically installed by default
  - Windows: Cygwin or Windows Subsystem for Linux (Windows 10)
- C compiler
  - Mac: Xcode
  - Linux: gcc
  - Windows: Rtools or Rtools40 (Windows 10)
- [NGCHMSupportFiles](https://github.com/MD-Anderson-Bioinformatics/NGCHMSupportFiles) R package
  - This package requires Java >= 11
  - This package contains a compiled .jar file and minified JavaScript file


## R Packages Installation

The NG-CHM R package can be installed from CRAN:

```{r, results='hide', message=FALSE, warning=FALSE, eval=FALSE}
install.packages('NGCHM')
```

The NGCHMSupportFiles can be installed from
[our R-Universe repository](https://md-anderson-bioinformatics.r-universe.dev/packages):


```{r, results='hide', message=FALSE, warning=FALSE, eval=FALSE}
install.packages('NGCHMSupportFiles', repos = c('https://md-anderson-bioinformatics.r-universe.dev', 'https://cloud.r-project.org'))
```

## Using the Docker Image

There is also a Docker image, [ngchm/rstudio-ngchm](https://hub.docker.com/r/ngchm/rstudio-ngchm), 
containing RStudio, the NGCHM-R package, and all needed dependencies 
available on Docker Hub.

The YouTube Video, [How to Create Next-Generation Clustered Heat Map in R Studio](https://www.youtube.com/watch?v=O42w5P3A1_8)
details usage of the Docker image. Further installation and usage information is available 
in this the [RStudio NGCHM GitHub](https://github.com/MD-Anderson-Bioinformatics/rstudio-ngchm) project.

